---
title: "p8105_hw2_mgg2153"
author: "mggn"
date: "9/25/2020"
output: github_document
---

## Hello and welcome to the second homework

The first order of business is to load the tidyverse!
```{r setup, include=FALSE}
library(tidyverse)
```


### Problem 2: NYC transit data

```{r nyc_transit}
transit_df = read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
  janitor::clean_names() %>%
  select(line, station_name, station_latitude, station_longitude, 
         route1:route11, entrance_type, entry, ada) %>%
  mutate(
    entry = recode(entry, "YES" = "TRUE", "NO" = "FALSE")
  )

print(transit_df)
#I want to retain:  line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance
  
```

#### About the NYC transit data

This data set is about New York City's subway system, particularly the subway 
exits and entrances. It is important to note that the dataset being described 
below is the original dataset but after some minor manipulation on my end. In the
code above, you can see that I:  
  -kept only certain variables using the retain function  
  -recoded the variable entry to have true/false values using mutate and recode  
  -changed the variable names so that they follow good style conventions (lower
  case, with underscores in place of spaces) using the clean_names function from
  the janitor library  

The dataset has **`r nrow(transit_df)`** observations and **`r ncol(transit_df)`** variables, 
which are: **`r names(transit_df)`**  




```{r route_pivot}

transit_df_tidy =
pivot_longer(
  transit_df,
    route1:route7,
    names_to = "route_num",
    values_to = "route"
  )

print(transit_df_tidy)
```


### Problem 3: FiveThirtyEight data

```{r 538_data}

pols_df = read_csv("./data/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(mon,into = c("year", "month", "day")) %>%
  mutate(
    month = month.abb[as.factor(month)]
  ) 

snp_df = read_csv("./data/snp.csv") %>%
  janitor::clean_names() %>%
  separate(date,into = c("day", "month", "year")) %>%
  mutate(
    month = month.abb[as.factor(month)]
  ) 
  
    

print(pols_df)
  
```
.
