---
title: "p8105_hw2_mgg2153"
author: "mggn"
date: "9/25/2020"
output: github_document
---

## Hello and welcome to the second homework

The first order of business is to load the tidyverse!
```{r setup, include=FALSE}
library(tidyverse)
```


### Problem 2: NYC transit data

```{r nyc_transit}
transit_df = read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
  janitor::clean_names() %>%
  select(line, station_name, station_latitude, station_longitude, 
         route1:route11, entrance_type, entry, ada) %>%
  mutate(
    entry = recode(entry, "YES" = "TRUE", "NO" = "FALSE")
  )

print(transit_df)
#I want to retain:  line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance
  
```

#### About the NYC transit data

This data set is about New York City's subway system, particularly the subway 
exits and entrances.

The penguins dataset has **`r nrow(transit_df)`** observations and **`r ncol(transit_df)`** variables, 
which are: **`r names(transit_df)`**

### Problem 3: FiveThirtyEight data

```{r 538_data}
month_df =
  tibble(
    month = 1:12,
    month_name = month.abb
  )

pols_df = read_csv("./data/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(mon,into = c("year", "month", "day")) %>%
  mutate(
    month = as.integer(month),
    year = as.integer(year),
    day = as.integer(day)
  )

left_join(pols_df, month_df,by = "month") %>%
  relocate(year, month_name) %>% 
  select(-month)

print(pols_df)
  
```
.
