---
title: "p8105_hw2_mgg2153"
author: "mggn"
date: "9/25/2020"
output: github_document
---

## Hello and welcome to the second homework

The first order of business is to load the tidyverse, which I did in a code chunk
but I am hiding the chunk from showing when the file it knit :-)
I also loaded the readxl package, as it is needed for the Mr. Trash Wheel dataset
```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
```

### Problem 1: Mr. Trash Wheel (done in class)

We will read in the precipitation data for 2017 and 2018
Note: file name in lecture is different than the file name used below
```{r}

trashwheel_df = 
	read_xlsx("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
		sheet = "Mr. Trash Wheel",
		range = cell_cols("A:N")) %>% 
	janitor::clean_names() %>% 
	drop_na(dumpster) %>% 
	mutate(
		sports_balls = round(sports_balls),
		sports_balls = as.integer(sports_balls)
	)

precip_2018 = 
	read_excel( "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
		sheet = "2018 Precipitation",
		skip = 1
	) %>% 
	janitor::clean_names() %>% 
	drop_na(month) %>% 
	mutate(year = 2018) %>% 
	relocate(year)

precip_2017 = 
	read_excel(
		"./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
		sheet = "2017 Precipitation",
		skip = 1
	) %>% 
	janitor::clean_names() %>% 
	drop_na(month) %>% 
	mutate(year = 2017) %>% 
	relocate(year)
```


Next, the two dataframes will be combined by the variable month. This will be 
possible through the creation of dataframe month_df, and then performing a left
join using the variable month

```{r}
month_df = 
	tibble(
		month = 1:12,
		month_name = month.name
	)

precip_df = 
	bind_rows(precip_2018, precip_2017)

precip_df =
	left_join(precip_df, month_df, by = "month")
```

#### About the Mr. Trashwheel dataset

Mr. Trashwheel is a "water vessel that removes trash from the Inner Harbor in
Baltimore, MD." Mr. Trashwheel has eyes! 

There are a total of **`r nrow(trashwheel_df)`** observations and **`r ncol(trashwheel_df)`**
variables, which are **`r names(trashwheel_df)`**

The median number of sports balls found in a dumpster in 2017 was **`r trashwheel_df %>% filter(year == 2017) %>% pull(sports_balls) %>% median()`**.

The total precipitation in 2018 was **`r precip_df %>% filter(year == 2018) %>% pull(total) %>% sum()`** inches.

### Problem 2: NYC transit data

First, I want to
(1) read in the data
(2) clean the variable names to follow good naming conventions
(3) keep only the variables I want
(4) recode entry to TRUE/FALSE instead of YES/NO

```{r nyc_transit}
transit_df = read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
  janitor::clean_names() %>%
  select(line, station_name, station_latitude, station_longitude, 
         route1:route11, entrance_type, entry, vending, ada) %>%
  mutate(
    entry = recode(entry, "YES" = "TRUE", "NO" = "FALSE")
  )
```

#### About the NYC transit data

This data set is about New York City's subway system, particularly the subway 
exits and entrances. It is important to note that the dataset being described 
below is the dataset that has been cleaned with the code above.

The dataset has **`r nrow(transit_df)`** observations and **`r ncol(transit_df)`** variables, 
which are: **`r names(transit_df)`**  

There are **`r transit_df %>% distinct(line, station_name) %>% count()`** distinct stations.

There are **`r transit_df %>% filter(ada == "TRUE") %>% count()`** stations that are ADA compliant.

And we can see that **`r transit_df %>% filter(vending == "NO" & entry == "TRUE") %>% count()`** stations without vending allow entry


After running these commands, Jeff asks us:  

#### are these data tidy?

To which I respond no! And Hadley Wickham would say "hell no"
One example of the lack of tidiness is that the routes, for example, are treated
as their own separate variables (ie, each route has its own column) when, to be
normalized, the routes should be collapsed into one variable. We can normalize/
"tidy" the data by using pivot_longer:

```{r route_pivot}

transit_df_tidy =
  transit_df %>%
  mutate(
    route8 = as.character(route8),
    route9 = as.character(route9),
    route10 = as.character(route10),
    route11 = as.character(route11)
    ) %>%
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route_name"
  )
```



With this "tidier" dataset, we can see that there are **`r transit_df_tidy %>% filter(route_name == "A") %>% distinct(line, route_name) %>% count()`** distinct A stations, and **`r transit_df_tidy %>% filter(route_name == 'A' & ada == "TRUE") %>% count()`** A stations
that are ADA compliant.


### Problem 3: FiveThirtyEight data

```{r 538_data}

pols_df = read_csv("./data/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(mon,into = c("year", "month", "day")) %>%
  mutate(
    month = month.abb[as.factor(month)]
  ) %>%
  mutate(
    president <- ifelse(prez_gop == 1, "gop", "dem") #need to take into account that in 1974, gop is 2; can change 2 to 1?
  ) %>%
  select(-prez_gop, -prez_dem, -day)
  

snp_df = read_csv("./data/snp.csv") %>%
  janitor::clean_names() %>%
  separate(date,into = c("day", "month", "year")) %>%
  mutate(
    month = month.abb[as.factor(month)],
    year = as.factor(year)
  ) %>%
  relocate(year,month)

unemployment_df = read_csv("./data/unemployment.csv") %>%
  mutate(
    Year = as.factor(Year)
  ) %>%
  pivot_longer(
    Jan:Dec,
    names_to = "month",
    values_to = "pct_unemployed"
  ) %>%
  rename(
    year = Year
  ) %>% mutate (
    month = month.abb[as.factor(month)]
  )
 
composite_538 =
  left_join(pols_df, snp_df, by = c("month", "year"))
            
composite_again =
  left_join(composite_538, unemployment_df, by = c("month", "year"))
```
.
